---
title: "Lab 6"
author: "Corinna Hong"
date: "February 13, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```

```{r}

library(tidyverse)
library(here)
library(janitor)
library(raster)
library(sf)
library(tmap)
library(tmaptools)
library(gstat)

```

# Grand Canyon GeoTIFF

```{r}

gc_dem <- raster(here("data","gc_dem.tif"))

```

```{r}

# Look at it with a base plot
plot(gc_dem)

# Check the CRS
gc_dem@crs # Shows CRS: WGS84

#Check extent
gc_dem@extent



# Create a WGS84 with lat and long metrics
wgs84 <- "+proj=longlat +zone=12 +datum=WGS84 +units=m +no_defs"

gc_reproj <- projectRaster(gc_dem, crs = wgs84, method = "bilinear")

gc_reproj@extent

```

### Crop raster to smaller area

```{r}

bounds <- as(extent(-112.4, -112.0, 36.1, 36.3), 'SpatialPolygons')

# assign same b projection as gc_proj
crs_bounds <- crs(gc_reproj)

# crop it
gc_crop <- crop(gc_reproj, bounds)

plot(gc_crop)

```

### Resample using the 'aggregate()' function

```{r}

gc_agg <- raster::aggregate(gc_crop, fact = 10) # calculates mean for 10x10 units of cells

plot(gc_agg)

```
 
### Plot in ggplot
 
```{r}

gc_df <- as.data.frame(gc_agg, xy = TRUE) # xy = true keeps spatial component, the xy are longlat values

ggplot(data = gc_df, aes(x = x, y = y)) +
  geom_raster(aes(fill = gc_dem)) +
  coord_quickmap() + # gives it a projection
  theme_minimal() +
  scale_fill_gradientn(colors = c("red",
                                  "orange",
                                  "yellow",
                                  "white"))


```

### Select cells that match given criteria
pretend there is a species of veg that will only grow when gc_dem is between 1000 to 1500

```{r}

gc_hab <- gc_crop

# set any cells outside of 1000 to 1500 to NA
gc_hab[gc_hab > 1500 | gc_hab < 1000] <- NA # Look for anything over 1500 and less than 1000 and assign them to NA

plot(gc_hab)

```

### Make it interactive with tmap
```{r, eval = FALSE}

tmap_mode("view")

# Error: no layer elements defined after tm_shape

tm_shape(gc_hab) + 
  tm_raster(legend.show = FALSE, palette = "plasma")

```

# Kriging rain in Kansas

```{r}

ks_counties <- read_sf(here("data","ks_counties","ks_counties_shapefile.shp"))

plot(ks_counties)

```

### Give it a projection
```{r}

# Check CRS:
st_crs(ks_counties) # hmmmm none...guess we should set one!

# Set to EPSG 4326 (WGS84 datum):
st_crs(ks_counties) <- 4326

# Now check again:
st_crs(ks_counties)

# And replot:
plot(ks_counties)

# That looks more like Kansas. 

```

### plot in ggplot

```{r}

ggplot(data = ks_counties) +
  geom_sf()

```

### read in rainfall data

```{r}

ks_rain <- read_csv(here("data", "ks_rain.csv")) %>% 
  clean_names()

```

### Update ks_rain to be recognized as spatial points

```{r}

ks_sf <- st_as_sf(ks_rain, coords = c("lon", "lat"), crs = 4326) # coords tell R which column is longitute and which is latitude

```

```{r}

ggplot() +
  geom_sf(data = ks_counties) +
  geom_sf(data = ks_sf,
          aes(color = amt, size = amt))

```

### Kriging to predict rainfall

```{r}

ks_sp <- as_Spatial(ks_sf)

# Check the class
class(ks_sf)

```

### Make a grid to make preditions over

```{r}

lat <- seq(37, 40, length.out = 200)
long <- seq(-94.6, -102, length.out = 200)


# turn this into a spatial grid

grid <- expand.grid(lon = long, lat = lat)
grid_sf <- st_as_sf(grid, coords = c("lon","lat"), crs = 4326)
grid_sp <- as_Spatial(grid_sf)

plot(grid_sp)

```

### Make variogram

```{r}

ks_vgm <- variogram(amt ~ 1, data = ks_sp)
plot(ks_vgm)

# Guesses for elemnets
# Nugget = 0.2
# Range = 200
# Sill = 0.8

ks_vgm_fit <- fit.variogram(ks_vgm, model = vgm(nugget = 0.2, psill = 0.8, model = "Sph", range = 200)) # sph is spherical model

plot(ks_vgm, ks_vgm_fit) # graph variogram and our model

```

### Krige

```{r}

ks_krige <- krige(amt ~ 1, ks_sp, grid_sp, model = ks_vgm_fit)

spplot(ks_krige, "var1.pred")

```

### Make df of predictions

```{r}

ks_df <- data.frame(ks_krige@data["var1.pred"],
                    ks_krige@data["var1.var"],
                    ks_krige@coords) %>% 
  rename(longitude = coords.x1,
         latitude = coords.x2)


# Convert this df to sf object

rain_sf <- st_as_sf(ks_df, coords = c("longitude","latitude"))
st_crs(rain_sf) <- 4326

ggplot(rain_sf) +
  geom_sf(aes(color = var1.pred))

```

### Crop to Kansas outline

```{r}

ks <- read_sf(dsn = here("data","states"),
              layer = "cb_2017_us_state_20m") %>%  
  dplyr::select(NAME) %>% # more than one select, specify that we want the one in dplyr
  filter(NAME == "Kansas") %>% 
  st_transform(crs = 4326)

plot(ks)

```

```{r}
# Find intersection of two

rain_sf_ks <- st_intersection(rain_sf, ks)

# Plot it

ggplot(rain_sf_ks) +
  geom_sf(aes(color = var1.pred)) +
  theme_minimal()

```

